Matroids from families of null hypotheses

### 2013-08-15 presentation/presentation.tex
Illustrated and fixed hierarchical theorem

\providecommand{\alert}[1]{\textbf{#1}}

\title{Matroids from families of null hypotheses}
\author{Willi Maurer, Florian Klinglmueller}
\date{\today}
\hypersetup{
  pdfkeywords={},
  pdfsubject={},
  pdfcreator={Emacs Org-mode version 7.8.11}}

@@

Testing hypotheses of no difference in pairwise comparisons of $m>3$ treatments (categories):\\
Shaffer (1986) : Modified Sequentially Rejective Multiple test procedures; Example: Modified Bonferroni-Holm
\begin{itemize}
  \item "Static":  After j rejections, test remaining hypotheses at level $\alpha/t_j$, $t_j$ = max number of true hypotheses if $j$ are false.
  \item "Dynamic": After rejecting $H_{i_1},...,H_{i_j}$, test remaining hypotheses at level $\alpha/t^*_j$,  $t^*_j$ = number of true hypotheses if $H_{i_1},...,H_{i_j}$ are false.
  \item Example: $m=4, n=6$ :
    \begin{itemize}
      \item Static: $t_0=6, t_1,t_2,t_3 = 3, t_4=2, t_5=1$
      \item Dynamic: if, e.g., $H_{\{1,2\}}$ and $H_{\{3,4\}}$ are rejected then only $H_{\{1,3\}}$ and $H_{\{2,3\}}$, or $H_{\{1,4\}}$ and $H_{\{2,3\}}$ can be simultaneously true, hence $t^*_2 = 2 < t_2 =3$ in this case.
       \end{itemize}
\end{itemize}
\end{frame}

@@

Consider $S$ and a collection of subsets $\mathscr{C}$ for which holds
\begin{enumerate}
\item $C_1 \neq C_2 \in \mathscr{C}$, then $C_1 \nsubseteq C_2$,
\item $C_1 \neq C_2 \in \mathscr{C}$, and $i \in C_1 \cap C_2$ then
   $\exists C_3 \subseteq (C_1 \cup C_2)$,
\end{enumerate}
then $\mathscr{C}$ is the set of circuits of a matroid $M$ on $S$

\input{cycle}

@@

\begin{frame}
\frametitle{Independence axioms}
\begin{block}{Independence}
  Consider $S$ and a collection of subsets $\mathscr{I}$ for which
  holds
  \begin{enumerate}
  \item $\emptyset \in \mathscr{I}$,
  \item $I \in \mathscr{I}$ and $J \subseteq I$ then $J \in
    \mathscr{I}$,
  \item $I,J \in \mathscr{I}$, $|I| = |J + 1|$ then $\exists j \in I
    \setminus J$ such that $J \cup i \in \mathscr{I}$,
  \end{enumerate}
  then $\mathscr{I}$ is the set of independent sets of a matroid $M$
  on $S$. This defines an independence relation $i \nsim_M I$ iff
  $I\cap i \in \mathscr{I}$. 
\end{block}

\begin{block}{Circuit}
  $C \subset S$ is a circuit iff $C \neq \mathscr{I}$ and for all $I
  \subset C$, $I \in \mathscr{I}$ (\ie minimally dependent set)
\end{block}
\end{frame}

@@

\begin{block}{Dependence}
  Consider $S$ and a relation $\sim_M$ on elements and subsets of $S$,
  then if for any $x,y_1,...,y_n,z_1,...,z_m$ holds
  \begin{enumerate}
  \item $y_i \sim \{y_1,...,y_n\}$ for all $i \leq n$
  \item if $n\geq 1$, $x \sim_M \{y_1,...,y_n\}$ and $x \nsim_M
    \{y_1,...,y_{n-1}$ then $y_n \sim \{y_1,...,y_{n-1},x\}$
  \item if $x \sim_M \{y_1,...,y_n\}$ and for all $i \leq n$ $y_i
    \sim_M \{z_1,...,z_m\}$ then $x \sim_M \{z_1,...,z_m\}$
  \end{enumerate}
  then $\sim_M$ defines a dependence relation of a matroid on $S$.
\end{block}

@@

\begin{frame}
\frametitle{Reminder: Graphical approaches}
\end{frame}

\begin{frame}
\frametitle{Hierarchical tests}
\end{frame}

>>>>

\providecommand{\alert}[1]{\textbf{#1}}
\usetheme{Novartis}
\doNotUseNovartisLogoOnAllSlides
\doNotUseNovartisLogoOnTitlePage

\title{Matroids from families of null hypotheses}
\author{Florian Klinglmueller, Willi Maurer}
\date{\today}
\hypersetup{
  pdfkeywords={},
  pdfsubject={},
  pdfcreator={Emacs Org-mode version 7.8.11}}

@@

Testing hypotheses of no difference in pairwise comparisons of $m>3$ treatments (categories):\\
\cite{shaffer1986modified}: Modified Sequentially Rejective Multiple test procedures; Example: Modified Bonferroni-Holm
\begin{itemize}
  \item "Static":  After j rejections, test remaining hypotheses at level $\alpha/t_j$, $t_j$ = max number of true hypotheses if $j$ are false.
  \item "Dynamic": After rejecting $H_{i_1},...,H_{i_j}$, test remaining hypotheses at level $\alpha/t^*_j$,  $t^*_j$ = number of true hypotheses if $H_{i_1},...,H_{i_j}$ are false.
  \item Example: $m=4, n=6$ :
    \begin{itemize}
      \item Static: $t_0=6, t_1,t_2,t_3 = 3, t_4=2, t_5=1$
      \item Dynamic: if, e.g., $H_{\{1,2\}}$ and $H_{\{3,4\}}$ are rejected then only $H_{\{1,3\}}$ and $H_{\{2,3\}}$, or $H_{\{1,4\}}$ and $H_{\{2,3\}}$ can be simultaneously true, hence $t^*_2 = 2 < t_2 =3$ in this case.
       \end{itemize}
\end{itemize}
\end{frame}

@@

\begin{tabular}{ll}
  \begin{minipage}{.4\textwidth}
    \input{cycle}
  \end{minipage} &
  \begin{minipage}{.6\textwidth}
    Consider $S$ and a collection of subsets $\mathscr{C}$ for which holds
    \begin{enumerate}
    \item $C_1 \neq C_2 \in \mathscr{C}$, then $C_1 \nsubseteq C_2$,
    \item $C_1 \neq C_2 \in \mathscr{C}$, and $i \in C_1 \cap C_2$ then
      $\exists C_3 \subseteq (C_1 \cup C_2)$,
    \end{enumerate}
    then $\mathscr{C}$ is the set of circuits of a matroid $M$ on $S$
  \end{minipage}
\end{tabular}

@@

\begin{frame}
\frametitle{Independence axioms}

\begin{tabular}{ll}
  \begin{minipage}{.4\textwidth}
  \only<1>{\input{independent}}
  \only<2>{\input{cycle}}
  \invisible<1>{\alert{\footnotesize minimally dependent set!}}
  \end{minipage} &
  \begin{minipage}{.6\textwidth}
  Consider $S$ and a collection of subsets $\mathscr{I}$ for which
  holds
  \begin{enumerate}
  \item $\emptyset \in \mathscr{I}$,
  \item $I \in \mathscr{I}$ and $J \subseteq I$ then $J \in
    \mathscr{I}$,
  \item $I,J \in \mathscr{I}$, $|I| = |J + 1|$ then $\exists j \in I
    \setminus J$ such that $J \cup i \in \mathscr{I}$,
  \end{enumerate}
  then $\mathscr{I}$ is the set of independent sets of a matroid $M$
  on $S$. This defines an independence relation $i \nsim_M I$ iff
  $I\cap i \in \mathscr{I}$.

  \end{minipage}
\end{tabular}

@@

\begin{tabular}{ll}
  \begin{minipage}{.4\textwidth}
    \input{dependent}
  \end{minipage} &
  \begin{minipage}{.6\textwidth}
  Relation $\sim_M$ on elements and subsets of $S$,
  if for any $x \in S$, and sets $Y,Z \subseteq S$ holds
  \begin{enumerate}
  \item $y \sim Y$ for all $y \in Y$
  \item if $x \sim_M Y$ and $x \nsim_M Y\setminus y$
    then $y \sim x \cup (Y \setminus y)$
  \item if $x \sim_M Y$ and for all $y \in Y$, $y \sim_M Z$ then $x \sim_M Z$
  \end{enumerate}
  then $\sim_M$ defines a matroid on $S$.
  \end{minipage}
\end{tabular}

@@

\begin{frame}
\frametitle{Reminder: Graphical approaches}

\begin{tabular}{ll}
  \begin{minipage}{.4\textwidth}
    \begin{tikzpicture}[remember picture,overlay]
      \node[yshift=-2.5cm,xshift=-1.4cm] at (current page.north west){\input{qvatest}};
    \end{tikzpicture}

  \end{minipage} &
  \begin{minipage}{.6\textwidth}
    \begin{itemize}
    \item Provides weights $w_{i,I}$ for each elementary hypotheses
      $H_i$ and each non-empty subset $I \subset S$
    \item Weights are used to define a closed test procedure of
      weighted intersection hypotheses tests
    \item We have $\sum_{i \in S} w_{i,I} \leq 1$ (conservative),
    \item and $w_{i,I} \leq w_{i,J}$ for $J \subseteq I$ (consonance).
    \end{itemize}
  \end{minipage}
\end{tabular}
\end{frame}

\begin{frame}
  \frametitle{Example: Hierarchical LSD}
\begin{tabular}{ll}
  \begin{minipage}{.4\textwidth}
    \input{lsd2.tex}
  \end{minipage} &
  \begin{minipage}{.6\textwidth}
    \begin{itemize}
    \item 3 Hypotheses:
      \begin{enumerate}
      \item Placebo vs. High Dose
      \item Placebo vs. Low Dose
      \item High Dose vs. Low Dose
      \end{enumerate}
    \item Prefixed sequence: $H_1 \rightarrow H_2 \rightarrow H_3$
    \item Test $H_1$ at level $\alpha$
    \item If we reject $H_1$ test $H_2$ and $H_3$ at level $\alpha$
      simultaneously
    \item Can we generalize this?
    \end{itemize}
  \end{minipage}
\end{tabular}

\begin{frame}
  \frametitle{Fixed sequence test of pairwise mean comparisons}
\begin{itemize}
\item {\bf Picture of hierarchical test hereabouts}
\end{itemize}

\begin{block}{Theorem}
  Assume that $M(\mathscr{H})$ neither contains loops nor
  parallels. Assume that hypotheses $H_1$ through $H_r$ have been
  rejected. Consider that hypotheses $H_{r+1}$ through $H_{r+t}$ are to
  be tested using at level $\alpha_r$ according to a prefixed order.
  If for some $1 \leq s \leq t$ holds
  \begin{equation}
    \label{eq:cond.hierarchical}
    \forall 1 \leq l < k \leq s: \; \exists i \leq r: i \sim \{l,k\},
  \end{equation}
  hypotheses $H_{r+1}$ through $H_{r+s}$ can be tested simultaneously
  at full level $\alpha_ra$.
\end{block}

\begin{frame}
  \frametitle{Example: Multiple doses}

\begin{tabular}{ll}
  \begin{minipage}{.4\textwidth}
    \begin{tikzpicture}[overlay]
      \node[yshift=70bp,xshift=-40bp] at (current page.south west){\input{lsd4.tex}};
    \end{tikzpicture}
  \end{minipage} &
  \begin{minipage}{.6\textwidth}
    \begin{tikzpicture}[remember picture,overlay]
      \node[yshift=100bp,xshift=160bp] at (current page.north west){\input{lsd4t.tex}};
    \end{tikzpicture}
  \end{minipage}
\end{tabular}
\end{frame}
